### 2021-03-14 lics2021-rebuttal-51.txt
another rebuttal version

All three reviewers see the paper in a positive light, and we thank
them for their feedback. The complaints of reviewer 1 concern only the
presentation, which can easily improved with all the helpful comments
of reviewer 1.

Below we give more detailed answers to the questions raised by
reviewers 1 and 2.

*********************************************************************
Answer to review 1:
===================

We are very grateful for the detailed comments, and we will happily
follow all suggestions. In particular, we will include an outline of
the paper, give more intuitive explanations and a broader overview of
the proof. Also the remark about equality is helpful.

> Question p.3: require that adjacent vertices to have different colors?

Indeed, in [18] there is this requirement, but it is not needed
because of the correctness criterion. Therefore we did not include it.

> Question p.3: can different vertices have the same label?

Yes.

> Question p.3: What about the atoms True and False in the fographs?

Indeed, in classical logic they are not needed for semantic
soundness and completeness, because

True = a v -a  and  False = a ^ -a

However, they are part of the language, and they also make the proof
systems simpler to use. Therefore we decided to include them.

> Question p.7: is \overline{S{A}} the complement of S{A} ?

Yes.

> Question p.10: A\cdot is variable renaming?

Yes.

> Question p.10: why "resource management"?

We see a combinatorial proof as a way to separate the linear part of a
proof (in the sense of linear logic) from the resource management of a
proof (in the sense that contraction and weakening duplicate and
delete resources). As this section is about contraction and weakening,
we chose this title. We agree that this needs to be better explained.

> Question p.11: "strong" homomorphism?

No, we do not have this condition. Note that this lemma is only about
weakening, and therefore the needed conditions are full and
injective. The next lemma is about contraction, where we need
surjectivity.

*********************************************************************
Answer to review 2:
===================

Question about the proof of Lemma 30:

Indeed, the structural equivalences concerning the quantifiers are not
discussed in [35]. But in our case, the equivalences concerning the
quantifiers are admissible in the system without the exists-rule. In
particular, in the example given in the review, a proof of

(forall x. A) \/ B

can be straightforwardly transformed into a proof of

forall x. (A \/ B)

(if x is vacuous in B)

We agree that this should be explained better.