### 2021-03-15 lics2021-rebuttal-51.txt
tweaked rebuttal

All three reviewers see the paper in a positive light, and we thank
them for their feedback. The complaints of reviewer 1 concern only the
presentation, which can easily improved with all the helpful comments
of reviewer 1.

Below we give more detailed answers to the questions raised by
reviewers 1 and 2.

*********************************************************************
Answer to review 1:
===================

We are very grateful for the detailed comments, and we will happily
follow all suggestions. We will include an outline of the paper, give
more intuitive explanations, and provide a broad overview of
the proof. The remark about equality is helpful.

> Question p.3: require that adjacent vertices to have different colors?

It was included in [18] because it is part of the standard definition of
colouring in graph theory. However, it is redundant, hence optional,
since it is implied by the preclusion of bimatchings: if vw is edge
between like-coloured vertices (v~w), then {v,w} is a bimatching.

> Question p.3: can different vertices have the same label?

Yes. (See the left of Fig. 1, for example.)

> Question p.3: What about the atoms True and False in the fographs?

Indeed, in classical logic they are not needed for semantic
soundness and completeness, because

True = a v -a  and  False = a ^ -a

However, they are part of the language, and they also make the proof
systems simpler to use. Therefore we kept them.

> Question p.7: is \overline{S{A}} the complement of S{A} ?

Yes.

> Question p.10: A\cdot is variable renaming?

Yes.

> Question p.10: why "resource management"?

We see a combinatorial proof as a way to separate the linear part of a
proof (in the sense of linear logic) from the resource management of a
proof (in the sense that contraction and weakening duplicate and
delete resources). As this section is about contraction and weakening,
we chose this title. We agree that this needs to be better explained.

> Question p.11: "strong" homomorphism?

No, we do not have this condition. This lemma is only about
weakening, so the requisite conditions are full and injective. The next
lemma is about contraction, where we need surjectivity.

*********************************************************************
Answer to review 2:
===================

Question about the proof of Lemma 30:

Indeed, the structural equivalences concerning the quantifiers are not
discussed in [35]. But in our case, the equivalences concerning the
quantifiers are admissible in the system without the exists-rule. In
particular, in the example given in the review, a proof of

(forall x. A) \/ B

can be straightforwardly transformed into a proof of

forall x. (A \/ B)

(if x is vacuous in B)

We agree that this should be explained better.